---
title: "MSDS7333 Case_Study_10 Spam verses Ham"
author: "Scott Anderwald, Chandon Nahor, Celia Taylor"
date: "11/23/2017"
output:
  word_document: default
  html_document: default
---
#Introduction:
The determination of spam versus ham is the basis behind case study 10.  Users of email are sometimes overburden with the amount of mail that they receive.  Along with the large amount of legitimate mail, known as ham, email users receive spam or unwanted email.  Machine learning algorithms, if properly deployed, can aid in the determination between spam and ham.  This case study uses Naive Bayes machine learning algorithm to determine which messages are potentially spam.
```{r setup, Include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
#Background:
Data for this case study comes from the following website: http://www.stat.berkeley.edu/~nolan/data/spam/SpamAssassinMessages.zip.
This is the website address associated with the book or .pdf by Deborah Nolan and Duncan Temple Lang, "Data Science in R: A Case Studies Approach to Computational Reasoning and Problem Solving".  The dataset originates from the http://spamassassin.apache.org/ website. Justin Mason created SpamAssassin in 1997.  In the summer of 2004, SpamAssassin became part of the Apache Software Foundation project. 


```{r installed packages, include=FALSE, echo=FALSE}
installed.packages()
sessionInfo()
```
#Methods:
In this case study the project team will utilize Naive Bayes for the spam filtering.  In this method, a Naive Bayes classifier of a particular feature in a class is unrelated to the presence of any other feature(www.analyticsvidhya.com/blog/2017/09/naive-bayes-explained/).  Naive Bayes, like other machine learning algorithms, has both strengths and weaknesses that need understanding for correct use.  One of the major strengths for Naive Bayes is that it is simple, fast, and effective. Another strength of Naive Bayes is that it requires few examples for training.  One weakness of Naive Bayes is that it may rely on faulty assumptions of equally important and independent features.  Additionally, Naive Bayes' estimated probabilities are less reliable than the predicted classes.  (Source Machine Learning with R Brett Lantz). 

For the case study the project team decided on answering question 19.  Question 19 requires adjusting parameters for the rpart(Recursive Partitioning and Regression Trees) package to develop a better model.

There are various parameters that control the rpart fit.  We primarily examine four parameters.

Parameter minsplit controls the minimum number of observations that must exist in a node in order for the split to be attempted.

Parameter minbucket controls the number of observations in any terminal note.  

Parameter cp is the complexity parameter which affects the overall fit.  

Parameter Maxcompete controls the number of competitor splits retained in the outputs. 

Information concerning the rpart.control package was obtained from the following website(https://www.rdocumentation.org/packages/rpart/versions/4.1-11/topics/rpart.control).   

To find the lowest Type I and Type II error, we executed fits = lapply code that tried parameter values above and below the default values that were in the base run. We first worked with the cp parameter until we found the lowest values for Type I and Type II error.  The primary goal was to lower the Type I error, but we also wanted to optimize the sum of the Type I and Type II error values.  Next we optimized minbucket.  We did try some runs that changed the other parameter values, but these runs did not give as much effect as changing cp and minbucket values.

```{r Load Packages, include=FALSE, echo=FALSE}

#load packages and libraries
library(knitr)
library(rmarkdown)
library(tm)
library(RColorBrewer)
library(rpart)
library(rpart.plot)
library(plotly)
```

```{r Directory Setup, include=FALSE, echo=TRUE}
spamPath = "~/Documents/ProjectFolder/R/Case_Studies/Case_Study_10"
setwd(spamPath)

list.dirs(spamPath, full.names = FALSE)

list.files(path = paste(spamPath, "SpamAssassinMessages", 
                        sep = .Platform$file.sep))

head(list.files(path = paste(spamPath, "SpamAssassinMessages", "spam_2",
                             sep = .Platform$file.sep)))

dirNames = list.files(path = paste(spamPath, "SpamAssassinMessages", 
                                   sep = .Platform$file.sep))
length(list.files(paste(spamPath, "SpamAssassinMessages", dirNames, 
                        sep = .Platform$file.sep)))

sapply(paste(spamPath, "SpamAssassinMessages", dirNames, 
             sep = .Platform$file.sep), 
       function(dir) length(list.files(dir)) )  # how many files are there

fullDirNames = paste(spamPath, "SpamAssassinMessages", dirNames, 
                     sep = .Platform$file.sep)

fileNames = list.files(fullDirNames[1], full.names = TRUE)
fileNames[1]

msg = readLines(fileNames[1])
head(msg)  # Too much output: how can you recursively traverse to get "short" output of the contents of an object

indx = c(1:5, 15, 27, 68, 69, 329, 404, 427, 516, 852, 971)
fn = list.files(fullDirNames[1], full.names = TRUE)[indx]
sampleEmail = sapply(fn, readLines)        

length(sampleEmail)
sapply(sampleEmail, length)   # always available!



msg = sampleEmail[[1]]
which(msg == "")[1]     # message body begins below first empty line; header above
# `which` returns all indices where condition is true
match("", msg)         # similar to `which`

splitPoint = match("", msg)

msg[ (splitPoint - 2):(splitPoint + 6) ] # look "around" split point  # Quiz question: is there an easier way to do this?

header = msg[1:(splitPoint-1)]
body = msg[ -(1:splitPoint) ]

splitMessage = function(msg) {       # split any message by empty line split point
  splitPoint = match("", msg)
  header = msg[1:(splitPoint-1)]
  body = msg[ -(1:splitPoint) ]
  return(list(header = header, body = body))
}

# split that 15 message sample
sampleSplit = lapply(sampleEmail, splitMessage)  # `list` is all about item isolation

# investigate a single header
header = sampleSplit[[1]]$header      
grep("Content-Type", header)

grep("multi", tolower(header[46]))    # "Content-Type: text/plain; charset=us-ascii" though contains "multi-part", if has attachment, etc.
#integer(0)   # another way of saying no result
#R syntax: an array of zero elements, expected type `integer`

header[46]

headerList = lapply(sampleSplit, function(msg) msg$header)
CTloc = sapply(headerList, grep, pattern = "Content-Type")   # locations of lines that contain "Content-Type"
CTloc   
# R moment: maintaining row/column names

# Bonus, quiz: how do you get the class/type of CTloc?  dim, class, length, head, sapply

# Rewrite previous function in way that collapses and avoids headers that have no Content-Type
sapply(headerList, function(header) {
  CTloc = grep("Content-Type", header)
  if (length(CTloc) == 0) return(NA)
  CTloc
})

hasAttach = sapply(headerList, function(header) {
  CTloc = grep("Content-Type", header)
  if (length(CTloc) == 0) return(FALSE)
  grepl("multi", tolower(header[CTloc])) 
})

hasAttach
sum(hasAttach) # 9 out of 15

```

```{r Begin boundary analysis remove attachments, include=FALSE, echo=FALSE}
header = sampleSplit[[6]]$header
boundaryIdx = grep("boundary=", header)
header[boundaryIdx]

sub(".*boundary=\"(.*)\";.*", "\\1", header[boundaryIdx])   # `sub`: replace first match with second argument, return new string

# regex's are "write-only"
# cf. the language coffeescript, which has a nice treatment

header2 = headerList[[9]]
boundaryIdx2 = grep("boundary=", header2)
header2[boundaryIdx2]

sub('.*boundary="(.*)";.*', "\\1", header2[boundaryIdx2])    # same regex doesn't work on this boundary, try different approach

boundary2 = gsub('"', "", header2[boundaryIdx2])   # get rid of quotation marks

sub(".*boundary= *(.*);?.*", "\\1", boundary2)

boundary = gsub('"', "", header[boundaryIdx])
sub(".*boundary= *(.*);?.*", "\\1", boundary)

sub(".*boundary= *([^;]*);?.*", "\\1", boundary)  # get rid of semicolon at the end
# bonus points: what change got rid of the semicolon and why?


getBoundary = function(header) {
  boundaryIdx = grep("boundary=", header)
  boundary = gsub('"', "", header[boundaryIdx])
  gsub(".*boundary= *([^;]*);?.*", "\\1", boundary)
}

#sampleSplit[[6]]$body

# Example from doc 15: get the boundary, then look for it in the body
boundary = getBoundary(headerList[[15]]) 
body = sampleSplit[[15]]$body

bString = paste("--", boundary, sep = "")
bStringLocs = which(bString == body)
bStringLocs        # shows up in two places, the two beginnings of each part

eString = paste("--", boundary, "--", sep = "")
eStringLoc = which(eString == body)
eStringLoc        # one ending

msg = body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1)]   #indexes look daunting: how to understand easily? exclusive interval
tail(msg)

att = c(msg, body[ (eStringLoc + 1) : length(body) ])   # to the "end", length(body)
tail(att)

# create function `dropAttach`, which gets rid of the second chunk (the attachment)
dropAttach = function(body, boundary){
  
  bString = paste("--", boundary, sep = "")
  bStringLocs = which(bString == body)
  
  if (length(bStringLocs) <= 1) return(body)
  
  eString = paste("--", boundary, "--", sep = "")
  eStringLoc = which(eString == body)
  if (length(eStringLoc) == 0) 
    return(body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1)])
  
  n = length(body)
  if (eStringLoc < n) 
     return( body[ c( (bStringLocs[1] + 1) : (bStringLocs[2] - 1), 
                    ( (eStringLoc + 1) : n )) ] )
  
  return( body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1) ])
}

```

```{r Extracing words, include=FALSE, echo=FALSE}
head(sampleSplit[[1]]$body)

msg = sampleSplit[[3]]$body
head(msg)

msg[ c(1, 3, 26, 27) ]

# memorize that :blank: is a useful part of a regex
cleanMsg = tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))   # replace all punctuation, numbers, "blanks" with space
cleanMsg[ c(1, 3, 26, 27) ]

# Drop common small words, aka stop words

stopWords = stopwords()
cleanSW = tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", stopWords))
SWords = unlist(strsplit(cleanSW, "[[:blank:]]+"))
SWords = SWords[ nchar(SWords) > 1 ]
stopWords = unique(SWords)
head(stopWords)


words = unlist(strsplit(cleanMsg, "[[:blank:]]+"))   # extract words from cleaned message separated by space
      # `unlist` is a very useful function

words = words[ nchar(words) > 1 ]        # get rid of single character words

words = words[ !( words %in% stopWords) ]   # get rid of stop words    #R idiom: %in%: quiz: how do get this functionality with other fns?
head(words)


# Extract into functions
cleanText = function(msg)   {
  tolower(gsub("[[:punct:]0-9[:space:][:blank:]]+", " ", msg))
}

findMsgWords = function(msg, stopWords) {
 if(is.null(msg))
  return(character())

 words = unique(unlist(strsplit(cleanText(msg), "[[:blank:]\t]+")))
 
 # drop empty and 1 letter words
 words = words[ nchar(words) > 1]
 words = words[ !( words %in% stopWords) ]
 invisible(words)
}

processAllWords = function(dirName, stopWords) {
       # read all files in the directory
  fileNames = list.files(dirName, full.names = TRUE)
       # drop files that are not email, i.e., cmds
  notEmail = grep("cmds$", fileNames)
  if ( length(notEmail) > 0) fileNames = fileNames[ - notEmail ]

  messages = lapply(fileNames, readLines, encoding = "latin1")
  
       # split header and body
  emailSplit = lapply(messages, splitMessage)
       # put body and header in own lists
  bodyList = lapply(emailSplit, function(msg) msg$body)
  headerList = lapply(emailSplit, function(msg) msg$header)
  rm(emailSplit)
  
       # determine which messages have attachments
  hasAttach = sapply(headerList, function(header) {
    CTloc = grep("Content-Type", header)
    if (length(CTloc) == 0) return(0)
    multi = grep("multi", tolower(header[CTloc])) 
    if (length(multi) == 0) return(0)
    multi
  })
  
  hasAttach = which(hasAttach > 0)
  
       # find boundary strings for messages with attachments
  boundaries = sapply(headerList[hasAttach], getBoundary)
  
       # drop attachments from message body
  bodyList[hasAttach] = mapply(dropAttach, bodyList[hasAttach], 
                               boundaries, SIMPLIFY = FALSE)
  
       # extract words from body
  msgWordsList = lapply(bodyList, findMsgWords, stopWords)
  
  invisible(msgWordsList)
}

msgWordsList = lapply(fullDirNames, processAllWords,       # process all the files: long compute time! (30s on mine)
                      stopWords = stopWords)      # notice you can put additional arguments in

numMsgs = sapply(msgWordsList, length)
numMsgs

isSpam = rep(c(FALSE, FALSE, FALSE, TRUE, TRUE), numMsgs)    # rep second argument is # of times: pairwise repetition
                                                          # straightforward idea; can you think of any other way to do this?

jjj = unlist(msgWordsList, recursive = FALSE)

msgWordsList = unlist(msgWordsList, recursive = FALSE)    # unlist only the first level down  # not idempotent! (why is this bad?)
   # Non-idempotent, stateful, no acknowledgement of completion, no easy way to backtrack


numEmail = length(isSpam)
numSpam = sum(isSpam)
(numHam = numEmail - numSpam)       # wrap in parens to see output

```

```{r Implemeting naive bayes classifier, include=FALSE, echo=FALSE}
set.seed(418910)      # `sample` is a random function: by using a set seed, we can reproduce our results during dev/debugging

# extract 1/3 as test data
testSpamIdx = sample(numSpam, size = floor(numSpam/3))   
testHamIdx = sample(numHam, size = floor(numHam/3))

testMsgWords = c((msgWordsList[isSpam])[testSpamIdx],
                 (msgWordsList[!isSpam])[testHamIdx] )
trainMsgWords = c((msgWordsList[isSpam])[ - testSpamIdx], 
                  (msgWordsList[!isSpam])[ - testHamIdx])

testIsSpam = rep(c(TRUE, FALSE), 
                 c(length(testSpamIdx), length(testHamIdx)))     # again pairwise rep
trainIsSpam = rep(c(TRUE, FALSE), 
                 c(numSpam - length(testSpamIdx), 
                   numHam - length(testHamIdx)))

```

```{r, include= FALSE, echo=TRUE}
bow = unique(unlist(trainMsgWords))     # bow: Bag of Words. BOW is a common (somewhat naïve) model in NLP

length(bow)

# Create an array called `spamWordCounts` of all the words and their respective frequencies in spam
spamWordCounts = rep(0, length(bow))
names(spamWordCounts) = bow       #name the buckets in the array

#cf. p130
tmp = lapply(trainMsgWords[trainIsSpam], unique) # get all the unique words per message, then unpack
tt = table( unlist(tmp) )                        # tabulate them (impossible w/o the computer)
spamWordCounts[ names(tt) ] = tt                 # Nice trick for filling in the spam array: index by names   ****

# Compute log likelihoods with this function:
computeFreqs = function(wordsList, spam, bow = unique(unlist(wordsList))) {
   # initialize a matrix for spam, ham, and log odds
  wordTable = matrix(0.5, nrow = 4, ncol = length(bow), 
                     dimnames = list(c("spam", "ham",              # store both probability and log odds
                                        "presentLogOdds", 
                                        "absentLogOdds"),  bow))

   # For each spam message, add 1 to counts for words in message
  counts.spam = table(unlist(lapply(wordsList[spam], unique)))
  wordTable["spam", names(counts.spam)] = counts.spam + .5         # Notice these .5 values: this is where "Naïve" Bayes comes from

   # Similarly for ham messages
  counts.ham = table(unlist(lapply(wordsList[!spam], unique)))  
  wordTable["ham", names(counts.ham)] = counts.ham + .5  


   # Find the total number of spam and ham
  numSpam = sum(spam)
  numHam = length(spam) - numSpam

   # Prob(word | spam) and Prob(word | ham)
  wordTable["spam", ] = wordTable["spam", ]/(numSpam + .5)
  wordTable["ham", ] = wordTable["ham", ]/(numHam + .5)
  
   # log odds
  wordTable["presentLogOdds", ] = 
     log(wordTable["spam",]) - log(wordTable["ham", ])
  wordTable["absentLogOdds", ] = 
     log((1 - wordTable["spam", ])) - log((1 -wordTable["ham", ]))

  invisible(wordTable)
}

trainTable = computeFreqs(trainMsgWords, trainIsSpam)
dim(trainTable)
```

```{r, include=FALSE, echo=TRUE}
newMsg = testMsgWords[[1]]

newMsg = newMsg[!is.na(match(newMsg, colnames(trainTable)))]

present = colnames(trainTable) %in% newMsg

sum(trainTable["presentLogOdds", present]) + 
  sum(trainTable["absentLogOdds", !present])

newMsg = testMsgWords[[ which(!testIsSpam)[1] ]]
newMsg = newMsg[!is.na(match(newMsg, colnames(trainTable)))]
present = (colnames(trainTable) %in% newMsg)
sum(trainTable["presentLogOdds", present]) + 
   sum(trainTable["absentLogOdds", !present])

computeMsgLLR = function(words, freqTable) {
       # Discards words not in training data.
  words = words[!is.na(match(words, colnames(freqTable)))]

       # Find which words are present
  present = colnames(freqTable) %in% words

  sum(freqTable["presentLogOdds", present]) +
    sum(freqTable["absentLogOdds", !present])
}

testLLR = sapply(testMsgWords, computeMsgLLR, trainTable)

tapply(testLLR, testIsSpam, summary)

pdf("SP_Boxplot.pdf", width = 6, height = 6)
spamLab = c("ham", "spam")[1 + testIsSpam]
boxplot(testLLR ~ spamLab, ylab = "Log Likelihood Ratio",
      #  main = "Log Likelihood Ratio for Randomly Chosen Test Messages",
        ylim=c(-500, 500))
dev.off()

typeIErrorRate = function(tau, llrVals, spam) {
  classify = llrVals > tau
  sum(classify & !spam)/sum(!spam)
}

typeIErrorRate(0, testLLR,testIsSpam)

typeIErrorRate(-20, testLLR,testIsSpam)

typeIErrorRates = function(llrVals, isSpam) {
  o = order(llrVals)
  llrVals =  llrVals[o]
  isSpam = isSpam[o]

  idx = which(!isSpam)
  N = length(idx)
  list(error = (N:1)/N, values = llrVals[idx])
}

typeIIErrorRates = function(llrVals, isSpam) {
  o = order(llrVals)
  llrVals =  llrVals[o]
  isSpam = isSpam[o]
  
  idx = which(isSpam)
  N = length(idx)
  list(error = (1:(N))/N, values = llrVals[idx])
}  

xI = typeIErrorRates(testLLR, testIsSpam)
xII = typeIIErrorRates(testLLR, testIsSpam)
tau01 = round(min(xI$values[xI$error <= 0.01]))
t2 = max(xII$error[ xII$values < tau01 ])

```

```{r , echo=FALSE}
pdf("LinePlotTypeI+IIErrors.pdf", width = 8, height = 6)


cols = brewer.pal(9, "Set1")[c(3, 4, 5)]
plot(xII$error ~ xII$values,  type = "l", col = cols[1], lwd = 3,
     xlim = c(-300, 250), ylim = c(0, 1),
     xlab = "Log Likelihood Ratio Values", ylab="Error Rate")
points(xI$error ~ xI$values, type = "l", col = cols[2], lwd = 3)
legend(x = 50, y = 0.4, fill = c(cols[2], cols[1]),
       legend = c("Classify Ham as Spam", 
                  "Classify Spam as Ham"), cex = 0.8,
       bty = "n")
abline(h=0.01, col ="grey", lwd = 3, lty = 2)
text(-250, 0.05, pos = 4, "Type I Error = 0.01", col = cols[2])

mtext(tau01, side = 1, line = 0.5, at = tau01, col = cols[3])
segments(x0 = tau01, y0 = -.50, x1 = tau01, y1 = t2, 
         lwd = 2, col = "grey")
text(tau01 + 20, 0.05, pos = 4,
     paste("Type II Error = ", round(t2, digits = 2)), 
     col = cols[1])

dev.off()

k = 5
numTrain = length(trainMsgWords)
partK = sample(numTrain)
tot = k * floor(numTrain/k)
partK = matrix(partK[1:tot], ncol = k)

testFoldOdds = NULL
for (i in 1:k) {
  foldIdx = partK[ , i]
  trainTabFold = computeFreqs(trainMsgWords[-foldIdx], trainIsSpam[-foldIdx])
  testFoldOdds = c(testFoldOdds, 
               sapply(trainMsgWords[ foldIdx ], computeMsgLLR, trainTabFold))
}

testFoldSpam = NULL
for (i in 1:k) {
  foldIdx = partK[ , i]
  testFoldSpam = c(testFoldSpam, trainIsSpam[foldIdx])
}

xFoldI = typeIErrorRates(testFoldOdds, testFoldSpam)
xFoldII = typeIIErrorRates(testFoldOdds, testFoldSpam)
tauFoldI = round(min(xFoldI$values[xFoldI$error <= 0.01]))
tFold2 = xFoldII$error[ xFoldII$values < tauFoldI ]
```

```{r  Begin large number analysis, include=FALSE, echo=FALSE}
smallNums = rep((1/2)^40, 2000000)
largeNum = 10000

print(sum(smallNums), digits = 20)

print(largeNum + sum(smallNums), digits = 20)

for (i in 1:length(smallNums)) {
  largeNum = largeNum + smallNums[i]
}
print(largeNum, digits = 20)

sampleSplit = lapply(sampleEmail, splitMessage)

header = sampleSplit[[1]]$header
header[1:12]

header[1] = sub("^From", "Top-From:", header[1])
header[1]

headerPieces = read.dcf(textConnection(header), all = TRUE)

headerPieces[, "Delivered-To"]


headerVec = unlist(headerPieces)
dupKeys = sapply(headerPieces, function(x) length(unlist(x)))
names(headerVec) = rep(colnames(headerPieces), dupKeys)

headerVec[ which(names(headerVec) == "Delivered-To") ]

length(headerVec)

length(unique(names(headerVec)))

processHeader = function(header) {
       # modify the first line to create a key:value pair
  header[1] = sub("^From", "Top-From:", header[1])
  
  headerMat = read.dcf(textConnection(header), all = TRUE)
  headerVec = unlist(headerMat)
  
  dupKeys = sapply(headerMat, function(x) length(unlist(x)))
  names(headerVec) = rep(colnames(headerMat), dupKeys)
  
  return(headerVec)
}

headerList = lapply(sampleSplit, 
                    function(msg) {
                      processHeader(msg$header)} )

contentTypes = sapply(headerList, function(header) 
                                        header["Content-Type"])
names(contentTypes) = NULL
contentTypes

hasAttach = grep("^ *multi", tolower(contentTypes))
hasAttach

boundaries = getBoundary(contentTypes[ hasAttach ])
boundaries

boundary = boundaries[9]
body = sampleSplit[[15]]$body

bString = paste("--", boundary, sep = "")
bStringLocs = which(bString == body)
bStringLocs

eString = paste("--", boundary, "--", sep = "")
eStringLoc = which(eString == body)
eStringLoc

diff(c(bStringLocs[-1], eStringLoc))

### This code has mistakes in it - and we fix them later!
processAttach = function(body, contentType){
  
  boundary = getBoundary(contentType)
  
  bString = paste("--", boundary, "$", sep = "")
  bStringLocs = grep(bString, body)
  
  eString = paste("--", boundary, "--$", sep = "")
  eStringLoc = grep(eString, body)
  
  n = length(body)
  
  if (length(eStringLoc) == 0) eStringLoc = n + 1
  if (length(bStringLocs) == 1) attachLocs = NULL
  else attachLocs = c(bStringLocs[-1],  eStringLoc)
  
  msg = body[ (bStringLocs[1] + 1) : min(n, (bStringLocs[2] - 1), 
                                         na.rm = TRUE)]
  
  if ( eStringLoc < n )
    msg = c(msg, body[ (eStringLoc + 1) : n ])
  
  if ( !is.null(attachLocs) ) {
    attachLens = diff(attachLocs, lag = 1) 
    attachTypes = mapply(function(begL, endL) {
      contentTypeLoc = grep("[Cc]ontent-[Tt]ype", body[ (begL + 1) : (endL - 1)])
      contentType = body[ begL + contentTypeLoc]
      contentType = gsub('"', "", contentType )
      MIMEType = sub(" *Content-Type: *([^;]*);?.*", "\\1", contentType)
      return(MIMEType)
    }, attachLocs[-length(attachLocs)], attachLocs[-1])
  }
  
  if (is.null(attachLocs)) return(list(body = msg, attachInfo = NULL) )
  else return(list(body = msg, 
                   attachDF = data.frame(aLen = attachLens, 
                                         aType = attachTypes,
                                         stringsAsFactors = FALSE)))                                
}

bodyList = lapply(sampleSplit, function(msg) msg$body)
attList = mapply(processAttach, bodyList[hasAttach], 
                 contentTypes[hasAttach], 
                 SIMPLIFY = FALSE)

lens = sapply(attList, function(processedA) 
                           processedA$attachDF$aLen)
head(lens)

attList[[2]]$attachDF

body = bodyList[hasAttach][[2]]
length(body)

body[35:45]

processAttach = function(body, contentType){

  n = length(body)
  boundary = getBoundary(contentType)
 
  bString = paste("--", boundary, sep = "")
  bStringLocs = which(bString == body)
  eString = paste("--", boundary, "--", sep = "")
  eStringLoc = which(eString == body)
  
  if (length(eStringLoc) == 0) eStringLoc = n
  if (length(bStringLocs) <= 1) {
    attachLocs = NULL
    msgLastLine = n
    if (length(bStringLocs) == 0) bStringLocs = 0
  } else {
    attachLocs = c(bStringLocs[ -1 ],  eStringLoc)
    msgLastLine = bStringLocs[2] - 1
  }
  
  msg = body[ (bStringLocs[1] + 1) : msgLastLine] 
  if ( eStringLoc < n )
    msg = c(msg, body[ (eStringLoc + 1) : n ])
  
  if ( !is.null(attachLocs) ) {
    attachLens = diff(attachLocs, lag = 1) 
    attachTypes = mapply(function(begL, endL) {
      CTloc = grep("^[Cc]ontent-[Tt]ype", body[ (begL + 1) : (endL - 1)])
      if ( length(CTloc) == 0 ) {
        MIMEType = NA
      } else {
        CTval = body[ begL + CTloc[1] ]
        CTval = gsub('"', "", CTval )
        MIMEType = sub(" *[Cc]ontent-[Tt]ype: *([^;]*);?.*", "\\1", CTval)   
      }
      return(MIMEType)
    }, attachLocs[-length(attachLocs)], attachLocs[-1])
  }
  
  if (is.null(attachLocs)) return(list(body = msg, attachDF = NULL) )
  return(list(body = msg, 
             attachDF = data.frame(aLen = attachLens, 
                                     aType = unlist(attachTypes),
                                     stringsAsFactors = FALSE)))                                
}                       

readEmail = function(dirName) {
       # retrieve the names of files in directory
  fileNames = list.files(dirName, full.names = TRUE)
       # drop files that are not email
  notEmail = grep("cmds$", fileNames)
  if ( length(notEmail) > 0) fileNames = fileNames[ - notEmail ]

       # read all files in the directory
  lapply(fileNames, readLines, encoding = "latin1")
}

processAllEmail = function(dirName, isSpam = FALSE)
{
       # read all files in the directory
  messages = readEmail(dirName)
  fileNames = names(messages)
  n = length(messages)
  
       # split header from body
  eSplit = lapply(messages, splitMessage)
  rm(messages)

       # process header as named character vector
  headerList = lapply(eSplit, function(msg) 
                                 processHeader(msg$header))
  
       # extract content-type key
  contentTypes = sapply(headerList, function(header) 
                                       header["Content-Type"])
  
       # extract the body
  bodyList = lapply(eSplit, function(msg) msg$body)
  rm(eSplit)

       # which email have attachments
  hasAttach = grep("^ *multi", tolower(contentTypes))

       # get summary stats for attachments and the shorter body
  attList = mapply(processAttach, bodyList[hasAttach], 
                   contentTypes[hasAttach], SIMPLIFY = FALSE)
  
  bodyList[hasAttach] = lapply(attList, function(attEl) 
                                           attEl$body)
 
  attachInfo = vector("list", length = n )
  attachInfo[ hasAttach ] = lapply(attList, 
                                  function(attEl) attEl$attachDF)
 
       # prepare return structure
  emailList = mapply(function(header, body, attach, isSpam) {
                       list(isSpam = isSpam, header = header, 
                            body = body, attach = attach)
                     },
                     headerList, bodyList, attachInfo, 
                     rep(isSpam, n), SIMPLIFY = FALSE )
  names(emailList) = fileNames
  
  invisible(emailList)
}

emailStruct = mapply(processAllEmail, fullDirNames,
                     isSpam = rep( c(FALSE, TRUE), 3:2))      
emailStruct = unlist(emailStruct, recursive = FALSE)

sampleStruct = emailStruct[ indx ]

save(emailStruct, file="emailXX.rda")

header = sampleStruct[[1]]$header
subject = header["Subject"]
els = strsplit(subject, "")
all(els %in% LETTERS)

testSubject = c("DEAR MADAME", "WINNER!", "")

els = strsplit(testSubject, "")
sapply(els, function(subject) all(subject %in% LETTERS))


gsub("[[:punct:] ]", "", testSubject)

gsub("[^[:alpha:]]", "", testSubject)

isYelling = function(msg) {
  if ( "Subject" %in% names(msg$header) ) {
     el = gsub("[^[:alpha:]]", "", msg$header["Subject"])
     if (nchar(el) > 0) 
        nchar(gsub("[A-Z]", "", el)) < 1
     else 
        FALSE
  } else 
    NA
}

perCaps =
function(msg)
{
  body = paste(msg$body, collapse = "")

       # Return NA if the body of the message is "empty"
  if(length(body) == 0 || nchar(body) == 0) return(NA)

       # Eliminate non-alpha characters
  body = gsub("[^[:alpha:]]", "", body)
  capText = gsub("[^A-Z]", "", body)
  100 * nchar(capText)/nchar(body)
}

sapply(sampleStruct, perCaps)

funcList = list( 
 isRe = function(msg) {
       "Subject" %in% names(msg$header) &&
         length(grep("^[ \t]*Re:", msg$header[["Subject"]])) > 0
  },
  numLines = function(msg) 
                length(msg$body),
  isYelling = function(msg) {
    if ( "Subject" %in% names(msg$header) ) {
       el = gsub("[^[:alpha:]]", "", msg$header["Subject"])
       if (nchar(el) > 0) 
         nchar(gsub("[A-Z]", "", el)) < 1
       else 
         FALSE
    }
    else NA
  },
  perCaps = function(msg) {
    body = paste(msg$body, collapse = "")

         # Return NA if the body of the message is "empty"
    if(length(body) == 0 || nchar(body) == 0) return(NA)

         # Eliminate non-alpha characters
    body = gsub("[^[:alpha:]]", "", body)
    capText = gsub("[^A-Z]", "", body)
    100 * nchar(capText)/nchar(body)
  }
)

lapply(funcList, function(func) 
                   sapply(sampleStruct, function(msg) func(msg)))

createDerivedDF =
function(email = emailStruct, operations = funcList, 
         verbose = FALSE)
{
  els = lapply(names(operations),
               function(id) {
                 if(verbose) print(id)
                 e = operations[[id]]
                 v = if(is.function(e)) 
                        sapply(email, e)
                      else 
                        sapply(email, function(msg) eval(e))
                 v
         })

   df = as.data.frame(els)
   names(df) = names(operations)
   invisible(df)
}

sampleDF = createDerivedDF(sampleStruct)
head(sampleDF)

funcList = list(
  isSpam =
    expression(msg$isSpam)
  ,
  isRe =
    function(msg) {
      # Can have a Fwd: Re:  ... but we are not looking for this here.
      # We may want to look at In-Reply-To field.
      "Subject" %in% names(msg$header) && 
        length(grep("^[ \t]*Re:", msg$header[["Subject"]])) > 0
    }
  ,
  numLines =
    function(msg) length(msg$body)
  ,
  bodyCharCt =
    function(msg)
      sum(nchar(msg$body))
  ,
  underscore =
    function(msg) {
      if(!"Reply-To" %in% names(msg$header))
        return(FALSE)
      
      txt <- msg$header[["Reply-To"]]
      length(grep("_", txt)) > 0  && 
        length(grep("[0-9A-Za-z]+", txt)) > 0
    }
  ,
  subExcCt = 
    function(msg) {
      x = msg$header["Subject"]
      if(length(x) == 0 || sum(nchar(x)) == 0 || is.na(x))
        return(NA)
      
      sum(nchar(gsub("[^!]","", x)))
    }
  ,
  subQuesCt =
    function(msg) {
      x = msg$header["Subject"]
      if(length(x) == 0 || sum(nchar(x)) == 0 || is.na(x))
        return(NA)
      
      sum(nchar(gsub("[^?]","", x)))
    }
  ,
  numAtt = 
    function(msg) {
      if (is.null(msg$attach)) return(0)
      else nrow(msg$attach)
    }
   
  ,
  priority =
    function(msg) {
      ans <- FALSE
      # Look for names X-Priority, Priority, X-Msmail-Priority
      # Look for high any where in the value
      ind = grep("priority", tolower(names(msg$header)))
      if (length(ind) > 0)  {
        ans <- length(grep("high", tolower(msg$header[ind]))) >0
      }
      ans
    }
  ,
  numRec =
    function(msg) {
      # unique or not.
      els = getMessageRecipients(msg$header)
      
      if(length(els) == 0)
        return(NA)
      
      # Split each line by ","  and in each of these elements, look for
      # the @ sign. This handles
      tmp = sapply(strsplit(els, ","), function(x) grep("@", x))
      sum(sapply(tmp, length))
    }
  ,
  perCaps =
    function(msg)
    {
      body = paste(msg$body, collapse = "")
      
      # Return NA if the body of the message is "empty"
      if(length(body) == 0 || nchar(body) == 0) return(NA)
      
      # Eliminate non-alpha characters and empty lines 
      body = gsub("[^[:alpha:]]", "", body)
      els = unlist(strsplit(body, ""))
      ctCap = sum(els %in% LETTERS)
      100 * ctCap / length(els)
    }
  ,
  isInReplyTo =
    function(msg)
    {
      "In-Reply-To" %in% names(msg$header)
    }
  ,
  sortedRec =
    function(msg)
    {
      ids = getMessageRecipients(msg$header)
      all(sort(ids) == ids)
    }
  ,
  subPunc =
    function(msg)
    {
      if("Subject" %in% names(msg$header)) {
        el = gsub("['/.:@-]", "", msg$header["Subject"])
        length(grep("[A-Za-z][[:punct:]]+[A-Za-z]", el)) > 0
      }
      else
        FALSE
    },
  hour =
    function(msg)
    {
      date = msg$header["Date"]
      if ( is.null(date) ) return(NA)
      # Need to handle that there may be only one digit in the hour
      locate = regexpr("[0-2]?[0-9]:[0-5][0-9]:[0-5][0-9]", date)
      
      if (locate < 0)
        locate = regexpr("[0-2]?[0-9]:[0-5][0-9]", date)
      if (locate < 0) return(NA)
      
      hour = substring(date, locate, locate+1)
      hour = as.numeric(gsub(":", "", hour))
      
      locate = regexpr("PM", date)
      if (locate > 0) hour = hour + 12
      
      locate = regexpr("[+-][0-2][0-9]00", date)
      if (locate < 0) offset = 0
      else offset = as.numeric(substring(date, locate, locate + 2))
      (hour - offset) %% 24
    }
  ,
  multipartText =
    function(msg)
    {
      if (is.null(msg$attach)) return(FALSE)
      numAtt = nrow(msg$attach)
      
      types = 
        length(grep("(html|plain|text)", msg$attach$aType)) > (numAtt/2)
    }
  ,
  hasImages =
    function(msg)
    {
      if (is.null(msg$attach)) return(FALSE)
      
      length(grep("^ *image", tolower(msg$attach$aType))) > 0
    }
  ,
  isPGPsigned =
    function(msg)
    {
      if (is.null(msg$attach)) return(FALSE)
      
      length(grep("pgp", tolower(msg$attach$aType))) > 0
    },
  perHTML =
    function(msg)
    {
      if(! ("Content-Type" %in% names(msg$header))) return(0)
      
      el = tolower(msg$header["Content-Type"]) 
      if (length(grep("html", el)) == 0) return(0)
      
      els = gsub("[[:space:]]", "", msg$body)
      totchar = sum(nchar(els))
      totplain = sum(nchar(gsub("<[^<]+>", "", els )))
      100 * (totchar - totplain)/totchar
    },
  subSpamWords =
    function(msg)
    {
      if("Subject" %in% names(msg$header))
        length(grep(paste(SpamCheckWords, collapse = "|"), 
                    tolower(msg$header["Subject"]))) > 0
      else
        NA
    }
  ,
  subBlanks =
    function(msg)
    {
      if("Subject" %in% names(msg$header)) {
        x = msg$header["Subject"]
        # should we count blank subject line as 0 or 1 or NA?
        if (nchar(x) == 1) return(0)
        else 100 *(1 - (nchar(gsub("[[:blank:]]", "", x))/nchar(x)))
      } else NA
    }
  ,
  noHost =
    function(msg)
    {
      # Or use partial matching.
      idx = pmatch("Message-", names(msg$header))
      
      if(is.na(idx)) return(NA)
      
      tmp = msg$header[idx]
      return(length(grep(".*@[^[:space:]]+", tmp)) ==  0)
    }
  ,
  numEnd =
    function(msg)
    {
      # If we just do a grep("[0-9]@",  )
      # we get matches on messages that have a From something like
      # " \"marty66@aol.com\" <synjan@ecis.com>"
      # and the marty66 is the "user's name" not the login
      # So we can be more precise if we want.
      x = names(msg$header)
      if ( !( "From" %in% x) ) return(NA)
      login = gsub("^.*<", "", msg$header["From"])
      if ( is.null(login) ) 
        login = gsub("^.*<", "", msg$header["X-From"])
      if ( is.null(login) ) return(NA)
      login = strsplit(login, "@")[[1]][1]
      length(grep("[0-9]+$", login)) > 0
    },
  isYelling =
    function(msg)
    {
      if ( "Subject" %in% names(msg$header) ) {
        el = gsub("[^[:alpha:]]", "", msg$header["Subject"])
        if (nchar(el) > 0) nchar(gsub("[A-Z]", "", el)) < 1
        else FALSE
      }
      else
        NA
    },
  forwards =
    function(msg)
    {
      x = msg$body
      if(length(x) == 0 || sum(nchar(x)) == 0)
        return(NA)
      
      ans = length(grep("^[[:space:]]*>", x))
      100 * ans / length(x)
    },
  isOrigMsg =
    function(msg)
    {
      x = msg$body
      if(length(x) == 0) return(NA)
      
      length(grep("^[^[:alpha:]]*original[^[:alpha:]]+message[^[:alpha:]]*$", 
                  tolower(x) ) ) > 0
    },
  isDear =
    function(msg)
    {
      x = msg$body
      if(length(x) == 0) return(NA)
      
      length(grep("^[[:blank:]]*dear +(sir|madam)\\>", 
                  tolower(x))) > 0
    },
  isWrote =
    function(msg)
    {
      x = msg$body
      if(length(x) == 0) return(NA)
      
      length(grep("(wrote|schrieb|ecrit|escribe):", tolower(x) )) > 0
    },
  avgWordLen =
    function(msg)
    {
      txt = paste(msg$body, collapse = " ")
      if(length(txt) == 0 || sum(nchar(txt)) == 0) return(0)
      
      txt = gsub("[^[:alpha:]]", " ", txt)
      words = unlist(strsplit(txt, "[[:blank:]]+"))
      wordLens = nchar(words)
      mean(wordLens[ wordLens > 0 ])
    }
  ,
  numDlr =
    function(msg)
    {
      x = paste(msg$body, collapse = "")
      if(length(x) == 0 || sum(nchar(x)) == 0)
        return(NA)
      
      nchar(gsub("[^$]","", x))
    }
)


SpamCheckWords =
  c("viagra", "pounds", "free", "weight", "guarantee", "million", 
    "dollars", "credit", "risk", "prescription", "generic", "drug",
    "financial", "save", "dollar", "erotic", "million", "barrister",
    "beneficiary", "easy", 
    "money back", "money", "credit card")


getMessageRecipients =
  function(header)
  {
    c(if("To" %in% names(header))  header[["To"]] else character(0),
      if("Cc" %in% names(header))  header[["Cc"]] else character(0),
      if("Bcc" %in% names(header)) header[["Bcc"]] else character(0)
    )
  }

emailDF = createDerivedDF(emailStruct)
dim(emailDF)
save(emailDF, file = "spamAssassinDerivedDF.rda")

load("~/Documents/ProjectFolder/R/Case_Studies/Case_Study_10/spamAssassinDerivedDF.rda")
dim(emailDF)

perCaps2 =
function(msg)
{
  body = paste(msg$body, collapse = "")

       # Return NA if the body of the message is "empty"
  if(length(body) == 0 || nchar(body) == 0) return(NA)

       # Eliminate non-alpha characters and empty lines 
  body = gsub("[^[:alpha:]]", "", body)
  els = unlist(strsplit(body, ""))
  ctCap = sum(els %in% LETTERS)
  100 * ctCap / length(els)
}

pC = sapply(emailStruct, perCaps)
pC2 = sapply(emailStruct, perCaps2)
identical(pC, pC2)

indNA = which(is.na(emailDF$subExcCt))

indNoSubject = which(sapply(emailStruct, 
                            function(msg) 
                              !("Subject" %in% names(msg$header))))

all(indNA == indNoSubject)

all(emailDF$bodyCharCt > emailDF$numLines)


x.at = c(1,10,100,1000,10000,100000)
y.at = c(1, 5, 10, 50, 100, 500, 5000)
nL = 1 + emailDF$numLines
nC = 1 + emailDF$bodyCharCt
pdf("ScatterPlotNumLinesNumChars.pdf", width = 6, height = 4.5)
plot(nL ~ nC, log = "xy", pch=".", xlim=c(1,100000), axes = FALSE,
     xlab = "Number of Characters", ylab = "Number of Lines")
box() 
axis(1, at = x.at, labels = formatC(x.at, digits = 0, format="d"))
axis(2, at = y.at, labels = formatC(y.at, digits = 0, format="d")) 
abline(a=0, b=1, col="red", lwd = 2)
dev.off()

pdf("SPAM_boxplotsPercentCaps.pdf", width = 5, height = 5)

percent = emailDF$perCaps
isSpamLabs = factor(emailDF$isSpam, labels = c("ham", "spam"))
boxplot(log(1 + percent) ~ isSpamLabs,
        ylab = "Percent Capitals (log)")

dev.off()

logPerCapsSpam = log(1 + emailDF$perCaps[ emailDF$isSpam ])
logPerCapsHam = log(1 + emailDF$perCaps[ !emailDF$isSpam ])

qqplot(logPerCapsSpam, logPerCapsHam, 
       xlab = "Regular Email", ylab = "Spam Email", 
       main = "Percentage of Capital Letters (log scale)",
       pch = 19, cex = 0.3)

pdf("SPAM_scatterplotPercentCapsTotChars.pdf", width = 8, height = 6)
 
colI = c("#4DAF4A80", "#984EA380")
logBodyCharCt = log(1 + emailDF$bodyCharCt)
logPerCaps = log(1 + emailDF$perCaps)
plot(logPerCaps ~ logBodyCharCt, xlab = "Total Characters (log)",
     ylab = "Percent Capitals (log)",
     col = colI[1 + emailDF$isSpam],
     xlim = c(2,12), pch = 19, cex = 0.5)

dev.off()

table(emailDF$numAtt, isSpamLabs)

pdf("SPAM_mosaicPlots.pdf", width = 8, height = 4)

oldPar = par(mfrow = c(1, 2), mar = c(1,1,1,1))

colM = c("#E41A1C80", "#377EB880")
isRe = factor(emailDF$isRe, labels = c("no Re:", "Re:"))
mosaicplot(table(isSpamLabs, isRe), main = "",
           xlab = "", ylab = "", color = colM)

fromNE = factor(emailDF$numEnd, labels = c("No #", "#"))
mosaicplot(table(isSpamLabs, fromNE), color = colM,
           main = "", xlab="", ylab = "")

par(oldPar)

dev.off()

#3.11 Fitting the rpart() Model to the email Data

setupRpart = function(data) {
  logicalVars = which(sapply(data, is.logical))
  facVars = lapply(data[ , logicalVars], 
                   function(x) {
                      x = as.factor(x)
                      levels(x) = c("F", "T")
                      x
                   })
  cbind(facVars, data[ , - logicalVars])
}

emailDFrp = setupRpart(emailDF)


set.seed(418910)
testSpamIdx = sample(numSpam, size = floor(numSpam/3))
testHamIdx = sample(numHam, size = floor(numHam/3))

testDF = 
  rbind( emailDFrp[ emailDFrp$isSpam == "T", ][testSpamIdx, ],
         emailDFrp[emailDFrp$isSpam == "F", ][testHamIdx, ] )
trainDF =
  rbind( emailDFrp[emailDFrp$isSpam == "T", ][-testSpamIdx, ], 
         emailDFrp[emailDFrp$isSpam == "F", ][-testHamIdx, ])

rpartFit = rpart(isSpam ~ ., data = trainDF, method = "class")


prp(rpartFit, extra = 1)


pdf("SPAM_rpartTree.pdf", width = 7, height = 7)

prp(rpartFit, extra = 1)
dev.off()

predictions = predict(rpartFit, 
       newdata = testDF[, names(testDF) != "isSpam"],
       type = "class")
 
predsForHam = predictions[ testDF$isSpam == "F" ]
summary(predsForHam)

sum(predsForHam == "T") / length(predsForHam)

predsForSpam = predictions[ testDF$isSpam == "T" ]
sum(predsForSpam == "F") / length(predsForSpam)
#Page 163
complexityVals = c(seq(0.00001, 0.0001, length=19),
                   seq(0.0001, 0.001, length=19), 
                   seq(0.001, 0.005, length=9),
                   seq(0.005, 0.01, length=9))

fits = lapply(complexityVals, function(x) {
         rpartObj = rpart(isSpam ~ ., data = trainDF,
                          method="class", 
                          control = rpart.control(cp=x) )
           
         predict(rpartObj, 
                 newdata = testDF[ , names(testDF) != "isSpam"],
                 type = "class")
        })


spam = testDF$isSpam == "T"
numSpam = sum(spam)
numHam = sum(!spam)
errs = sapply(fits, function(preds) {
                      typeI = sum(preds[ !spam ] == "T") / numHam
                      typeII = sum(preds[ spam ] == "F") / numSpam
                      c(typeI = typeI, typeII = typeII)
                     })

pdf("SPAM_rpartTypeIandII.pdf", width = 8, height = 7)

cols = brewer.pal(9, "Set1")[c(3, 4, 5)]
plot(errs[1,] ~ complexityVals, type="l", col=cols[2], 
     lwd = 2, ylim = c(0,0.2), xlim = c(0,0.005), 
     ylab="Error", xlab="complexity parameter values")
points(errs[2,] ~ complexityVals, type="l", col=cols[1], lwd = 2)

text(x =c(0.003, 0.0035), y = c(0.12, 0.05), 
     labels=c("Type II Error", "Type I Error"))

minI = which(errs[1,] == min(errs[1,]))[1]
abline(v = complexityVals[minI], col ="grey", lty =3, lwd=2)

text(0.0007, errs[1, minI]+0.01, 
     formatC(errs[1, minI], digits = 2))
text(0.0007, errs[2, minI]+0.01, 
     formatC(errs[2, minI], digits = 3))

dev.off()


```
#Base Analysis:

The original run used parameters determined by the algorithm. The complexity parameter used a value of 0.001. Type I Error was 3.4% and Type II error was 13%.

```{r , echo=FALSE }
fits = lapply(complexityVals, function(x) {
         rpartObj = rpart(isSpam ~ ., data = trainDF,
                          method="class", 
                          control = rpart.control(minsplit =3,
                                                  cp = 0.01,
                                                  minbucket = 10))#added this to code compare line 1360
           
         predict(rpartObj, 
                 newdata = testDF[ , names(testDF) != "isSpam"],
                 type = "class")
        })



spam = testDF$isSpam == "T"
numSpam = sum(spam)
numHam = sum(!spam)
errs = sapply(fits, function(preds) {
                      typeI = sum(preds[ !spam ] == "T") / numHam
                      typeII = sum(preds[ spam ] == "F") / numSpam
                      c(typeI = typeI, typeII = typeII)
                     })

pdf("SPAM_rpartTypeIandII_Question19A.pdf", width = 8, height = 7)

cols = brewer.pal(9, "Set1")[c(3, 4, 5)]
plot(errs[1,] ~ complexityVals, type="l", col=cols[2], 
     lwd = 2, ylim = c(0,0.2), xlim = c(0,0.005), 
     ylab="Error", xlab="minsplit =3, cp = 0.01, minbucket = 10 ")
points(errs[2,] ~ complexityVals, type="l", col=cols[1], lwd = 2)

text(x =c(0.003, 0.0035), y = c(0.12, 0.05), 
     labels=c("Type II Error", "Type I Error"))

minI = which(errs[1,] == min(errs[1,]))[1]
abline(v = complexityVals[minI], col ="grey", lty =3, lwd=2)

text(0.0007, errs[1, minI]+0.01, 
     formatC(errs[1, minI], digits = 2))
text(0.0007, errs[2, minI]+0.01, 
     formatC(errs[2, minI], digits = 3))

dev.off()
```
#Analysis with Parameter Adjustment:
Run 1: minsplit =3 cp = 0.01, minbucket = 10
Run 2: minsplit =3 cp = 0.001, minbucket = 10
Run 3: minsplit =3 cp = 0.001, minbucket = 5
#Results:
Results from the base run for the type I and II errors plot shows a that a type I error is minimized at a complexity parameter value of 0.001.  This results are a 3.4% Type I error and 13.0% Type II error.  The complexity parameter is a mechanism for specifying the threshold for choosing a split for a subgroup. For this case study and Question 19, the parameters for rpart.control were adjusted. The parameters that were used for the projects were minsplit = 3, minbucket = 22, cp = 0.001, and maxsurrogate = 5.  The output for the was a 4.8% error for type I and 13% error rate for type II error. 


#Conclusions:
After comparing the base run to the runs executed for Question 19, it appears that changing additional rpart.control parameters' values slightly increases the accuracy as given in the decrease of the sum of Type I and Type II error.  The decrease realized was from Type I Error at 3.4% and Type II Error at 13.0% (sum 16.4%) to Type I Error at 3.4% and Type II Error at 11.6% (sum 15.0%).  There could be a time cost to the analysis which makes the slight increase in accuracy unreasonable.


```{r , echo=FALSE }
fits = lapply(complexityVals, function(x) {
         rpartObj = rpart(isSpam ~ ., data = trainDF,
                          method="class", 
                          control = rpart.control(minsplit =3,
                                                  minbucket = 10,
                                                  cp = 0.001))#added this to code compare line 1360
           
         predict(rpartObj, 
                 newdata = testDF[ , names(testDF) != "isSpam"],
                 type = "class")
        })



spam = testDF$isSpam == "T"
numSpam = sum(spam)
numHam = sum(!spam)
errs = sapply(fits, function(preds) {
                      typeI = sum(preds[ !spam ] == "T") / numHam
                      typeII = sum(preds[ spam ] == "F") / numSpam
                      c(typeI = typeI, typeII = typeII)
                     })

pdf("SPAM_rpartTypeIandII_Question19B.pdf", width = 8, height = 7)

cols = brewer.pal(9, "Set1")[c(3, 4, 5)]
plot(errs[1,] ~ complexityVals, type="l", col=cols[2], 
     lwd = 2, ylim = c(0,0.2), xlim = c(0,0.005), 
     ylab="Error", xlab="minsplit =3,cp = 0.001 minbucket = 10")
points(errs[2,] ~ complexityVals, type="l", col=cols[1], lwd = 2)

text(x =c(0.003, 0.0035), y = c(0.12, 0.05), 
     labels=c("Type II Error", "Type I Error"))

minI = which(errs[1,] == min(errs[1,]))[1]
abline(v = complexityVals[minI], col ="grey", lty =3, lwd=2)

text(0.0007, errs[1, minI]+0.01, 
     formatC(errs[1, minI], digits = 2))
text(0.0007, errs[2, minI]+0.01, 
     formatC(errs[2, minI], digits = 3))

dev.off()
```
```{r , echo=FALSE }
fits = lapply(complexityVals, function(x) {
         rpartObj = rpart(isSpam ~ ., data = trainDF,
                          method="class", 
                          control = rpart.control(minsplit =3,
                                                  minbucket = 5,
                                                  cp = 0.001))#added this to code compare line 1360
           
         predict(rpartObj, 
                 newdata = testDF[ , names(testDF) != "isSpam"],
                 type = "class")
        })



spam = testDF$isSpam == "T"
numSpam = sum(spam)
numHam = sum(!spam)
errs = sapply(fits, function(preds) {
                      typeI = sum(preds[ !spam ] == "T") / numHam
                      typeII = sum(preds[ spam ] == "F") / numSpam
                      c(typeI = typeI, typeII = typeII)
                     })

pdf("SPAM_rpartTypeIandII_Question19C.pdf", width = 8, height = 7)

cols = brewer.pal(9, "Set1")[c(3, 4, 5)]
plot(errs[1,] ~ complexityVals, type="l", col=cols[2], 
     lwd = 2, ylim = c(0,0.2), xlim = c(0,0.005), 
     ylab="Error", xlab="minsplit =3, cp = 0.001, minbucket = 5")
points(errs[2,] ~ complexityVals, type="l", col=cols[1], lwd = 2)

text(x =c(0.003, 0.0035), y = c(0.12, 0.05), 
     labels=c("Type II Error", "Type I Error"))

minI = which(errs[1,] == min(errs[1,]))[1]
abline(v = complexityVals[minI], col ="grey", lty =3, lwd=2)

text(0.0007, errs[1, minI]+0.01, 
     formatC(errs[1, minI], digits = 2))
text(0.0007, errs[2, minI]+0.01, 
     formatC(errs[2, minI], digits = 3))

dev.off()
```
```

