---
title: "Case_Study_12 Simulation Study of a Branching Process"
author: "Scott Anderwald, Chandon Nahar, Celia Taylor"
date: "12/9/2017"
output:
  word_document: default
  html_document: default
---
#Introduction:

In this assignment, we are trying to model the branching process using the Monte Carlo Method, which is just a way to generate a sample from a random distribution. The basic idea is that we start with a job which spawns other jobs. Upon job creation, that job can generate offspring jobs. The “parent” job takes a certain amount of time to finish. While this “parent” job is running, it can spawn other “child” jobs which cannot run until the parent job finishes.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Background:

This branching process is representative of a natural process as well and happens in computer programming. In addition, there is a recursive nature to this process. Therefore, the usual way of applying the probability distribution does not work. In essence, we will be generating these random jobs to fill in a high-dimensional space and our intent is to understand what the ultimate distribution turns out to be. 

This case study is from Chapter 7, "Simulation Study of a Branching Process", of the book or .pdf by Deborah Nolan and Duncan Temple Lang, "Data Science in R: A Case Studies Approach to Computational Reasoning and Problem Solving".

http://rdatasciencecases.org/ is the website address associated with this book.  Code associated with Chapter 7 is at http://rdatasciencecases.org/BirthDeathProcess/code.R. Data for this case study is generated randomly by the code.
 
```{r installed packages, include=FALSE, echo=FALSE}
installed.packages()
sessionInfo()
```
#Method:

In this case study, the project team will utilize a Monte Carlo Method to generate independent random results from its probability distribution.  Then we use the properties (mean and Standard Deviation) of the observed results as approximations to the expected properties.   Basically, we study the behavior of thousands of random variables in order to discern the probable behavior of the outcome.
The strengths of this process are that we can figure out the samples size to provide proper precision, use statistical principles to summarize and evaluate the results, and utilize experimental design to change parameters to better understand and study the process.
The cautionary aspects are that the study needs to be carefully designed so that the problem is actually reproducible and representative.  The results will not be the same every time because it is a random process, but will be reproducible by remaining statistically coherent.  Also, because the results are not the same every time, testing and debugging of the code must be done very carefully.
For this case study, the project team decided on answering Chapter 7, Question 10.  Question 10 requires executing the simulation study with different values for the parameters.  Question 10 suggests holding the kappa parameter constant at the value of 1 and running the simulation with various values of lambda.  
Parameter lambda is the birth rate.
Parameter kappa is the completion rate or the lifetime.
In order to keep the run time of the algorithm at a reasonable level, the number of random outcomes studied was reduced from 400 to 40.
The relationship between lambda and kappa is very important to how long the process runs and how many generations occur.  The simulation studies what happens when lambda/kappa ranges as less than one, close to one, and greater than one.  To do this, we ran 5 different trials with kappa remaining constant at one.
Values below original or baseline run
Trial 1 for Plot 1
kappa = 1 and lambda ranges from 0.01 to 0.2 by increments of 0.1
kappa = 1 and lambda ranges from 0.4 to 1.6 by increments of  0.2
kappa = 1 and lambda ranges from 1.25 to 2.0 by increments of  0.25

---------------
Trial 2 for Plot 2
kappa = 1 and lambda ranges from 0.0015 to 0.4 by increments of  0.1
kappa = 1 and lambda ranges from 0.6 to 1.6 by increments of  0.2
kappa = 1 and lambda ranges from 1.75 to 2.75 by increments of  0.25

---------------
Original or baseline run
Trial 3 for Plot 3
kappa = 1 and lambda ranges from 0.1 to 0.6 by increments of  0.1
kappa = 1 and lambda ranges from 0.8 to 2.0 by increments of  0.2
kappa = 1 and lambda ranges from 2.25 to 3.0 by increments of  0.25

---------------
Values above original or baseline run
Trial 4 for Plot 4
kappa = 1 and lambda ranges from 0.2 to 0.8 by increments of  0.1
kappa = 1 and lambda ranges from 1.0 to 2.2 by increments of  0.2
kappa = 1 and lambda ranges from 2.50 to 3.25 by increments of  0.25

---------------
Trial 5 for Plot 5
kappa = 1 and lambda ranges from 0.3 to 0.9 by increments of  0.1
kappa = 1 and lambda ranges from 1.6 to 2.8 by increments of  0.2
kappa = 1 and lambda ranges from 3.0 to 4.25 by increments of  0.25 


```{r Load Packages, include=FALSE, echo=FALSE}
library(knitr)
library(rmarkdown)
library(scatterplot3d)
```

```{r Directory Setup, include=FALSE, echo=FALSE}
#For the mac user
setwd("~/Documents/ProjectFolder/R/Case_Studies/Case_Study_12_Chap7")
getwd()

```


```{r include=FALSE, echo=FALSE}

empirical = replicate(6000, sum(rexp(n = 3)))

mean(empirical)

sd(empirical)

sum(empirical <= 5)/length(empirical)

rates = c(seq(0.1, 1, by = 0.1), seq(2, 7, by = 1))

samples = lapply(rates, function(r) {
             replicate(6000, sum(rexp(n = 3, rate = r))) })


pdf("BA_MCexample.pdf", width = 8, height = 5)
par(mar = c(2,4,1,1))
plot(0, 0, xlim = c(-0.1, 10), ylim = c(0, 1), pch = ".",
     xlab="", ylab="Cumulative Distribution")
xx = seq(-0.1, 20, by = 0.05)

invisible(lapply(samples[-(19:20)], function(s) {
  Fn = ecdf(s)                                         #empirical CDF function
  ptile = Fn(xx)                                       # apply empirical cdf to x values
  points(x = xx, y = ptile, type = "l")
}))

labs = c(1, 2, 3, 6, 10, 11, 16)
xs = c(8,8,8,6, 4, 2, 0.75)
poss = c(3,3,3, 2, 2, 2,2)
mapply(function(lab, s, x, r, p) 
  {
  Fn = ecdf(s)
  text(x = x, y = Fn(x), labels = r, pos = p, cex = 0.8)                  #location of the labels; note having to compute this
}, lab = labs, s = samples[labs], x = xs, r = rates[labs], p = poss)

dev.off()

```

```{r include=FALSE, echo=FALSE}
seedx = 116201412
set.seed(seedx)
# Note: we have to run these in order!


kappa = 0.3
# first job's lifetime
d0 = rexp(1, rate = kappa)
d0

lambda = 0.5
# first birth time
birth1 = rexp(1, rate = lambda)
birth1

# simulate next birth time
itime = c(birth1, rexp(1, rate = lambda))
itime

# simulate next one
itime = c(itime, rexp(1, rate = lambda))
itime

cumsum(itime)

itime = c(itime, rexp(1, rate = lambda))
btime = cumsum(itime)
btime # birth times

btime = btime[ btime < d0 ]      # only two were generated during first job's lifetime (< d0)

dtime = d0 + rexp(n = length(btime), rate = kappa) 
dtime  # birth times plus exp(kappa): death times: all start (in parallel after d0 finishes)

# these can potentially generate offspring as well: "birth and assassination process"
btime[1] + rexp(1, rate = lambda) # 6.233543

dtime[1] + rexp(1, rate = kappa)  # 12.55685
```

```{r include=FALSE, echo=FALSE}
# note defining signature of function great way to get started!
genKids = function(bTime, cTime, lambda = 0.5, kappa = 0.3) {    # four parameters in function definition 
         # Parent job born at bTime and completes at cTime
    
         # Birth time of first child
    mostRecent = rexp(1, rate = lambda) + bTime
    kidBirths = numeric()

    while (mostRecent < cTime) {
      kidBirths = c(kidBirths, mostRecent)
      mostRecent = mostRecent  + rexp(1, rate = lambda) 
    }
    
         # generate lifetimes for all offspring
    numKids = length(kidBirths)
    runtime = rexp(numKids, rate = kappa)
    kidCompletes = rep(cTime, numKids) + runtime
    
    data.frame(births = kidBirths, 
               completes = kidCompletes)
}

set.seed(seedx)   # reset the seed in case we lost track

genKids(1, 6)

genKids(1, 6)

genKids(1, 6)    # up arrow + return to get a feel for how these are "distributed"

# we can this recursively:
genBirth = function(currentTime, cTime, 
                    births = numeric(), lambda = 0.5) {
          
         # Generate birth time of next job after currentTime
    mostRecent = rexp(1, rate = lambda) + currentTime
    
    if (mostRecent > cTime)        # note here we "discard" generated data if it doesn't meet the condition
      return(births)
    else {
      births = c(births, mostRecent)
      genBirth(currentTime = mostRecent, cTime, births, lambda)     # recursive call
    }
  }

# Effectively, the recursive call replaces the `while` loop
# Deep CS concepts here of "tail-recursion" and convergence

# gen kids recursively:
genKidsR = function(bTime, cTime, lambda = 0.5, kappa = 0.3) {
         # Parent job born at bTime and completes at cTime
    
    kidBirths = genBirth(bTime, cTime, lambda = lambda)
    
         # generate lifetimes for all offspring
    numKids = length(kidBirths)
    runtime = rexp(numKids, rate = kappa)
    kidDeaths = rep(cTime, numKids) + runtime
    
    data.frame(births = kidBirths, 
               completes = kidDeaths)
} 

set.seed(seedx)   # reset again

numKids = replicate(1000, nrow(genKids(1, 6)))
mean(numKids)

eprobs = table(numKids)/length(numKids)
probs = dpois(x = 0:max(numKids), lambda = 2.5)

plot(eprobs, type = "h",
     ylab = "Proportion", xlab = "Number of offspring")
segments(x0 = 0.1 + 0:max(numKids), y0 = rep(0, max(numKids)), # nice plotting function: `segments`; goes along with type "h" + stagger by .1
         y1 = probs, col="grey", lwd = 2)
```

```{r include=FALSE, echo=FALSE}
# pdf("numSimOffspring.pdf", width = 8, height = 5)
# oldPar = par(mar = c(4.1, 4.1, 1, 1))
# 
# plot(eprobs, type = "h",
#      ylab = "Proportion", xlab = "Number of offspring")
# segments(x0 = 0.1 + 0:max(numKids), 
#          y0 = rep(0, max(numKids)), y1 = probs,
#          col="grey", lwd = 2)
# 
# par(oldPar)
# dev.off()

# compute the observed proportions vs. theoretical ones
obsCt = table(numKids)
expCt = 1000* c(dpois(0:8, lambda = 2.5), 
                ppois(8, lower.tail = FALSE, lambda = 2.5))

# Compute chi^2 statistic: recall! good for comparing frequencies in buckets
stat = sum((obsCt - expCt)^2/expCt)
pchisq(stat, df = 9, lower.tail = FALSE)  # 0.9943803 # p-value



# Rewrite above functionality returning a vector, so we can use vector comprehension next (that is, applying a function across a vector)
genKidsU =  function(bTime, cTime, lambda = 0.5, kappa = 0.3) {
         # Generate the birth times and assassination times
         # for the children of a job who is born at bTime 
         # and completed at cTime.
    lambda = (cTime - bTime) * lambda
    numKids = rpois(1, lambda = lambda)
    kidBirths = sort(runif(numKids, min = bTime, max = cTime))
    
         # generate lifetimes for each offspring
    runtime = rexp(numKids, rate = kappa)
    kidDeaths = rep(cTime, numKids) + runtime
    
    return(data.frame(births = kidBirths, completes = kidDeaths))
  }
```

```{r include=FALSE, echo=FALSE}
seedx = 116201412
set.seed(seedx)

time1 = system.time( replicate(4000, genKids(1, cTime = 9)) )
time2 = system.time( replicate(4000, genKidsU(1, cTime = 9)) )

time1/time2

time1 = system.time( replicate(4000, genKids(1, cTime = 100)) )
time2 = system.time( replicate(4000, genKidsU(1, cTime = 100)) )
time1/time2

set.seed(seedx)
Rprof("profGenKids1.out")
invisible( replicate(1000, genKids(1, cTime = 100)) )
Rprof(NULL)

summaryRprof("profGenKids1.out")$by.self

set.seed(seedx)
Rprof("profGenKidsU.out")
invisible( replicate(1000, genKidsU(1, cTime = 100)) )
Rprof(NULL)
summaryRprof("profGenKidsU.out")$by.self
```

```{r include=FALSE, echo=FALSE}
# apply genKidsU across a vector using `mapply`
genKidsV = function(bTimes, cTimes, lambda = 0.5, kappa = 0.3) {
       # bTimes & cTimes - vector of birth and completion times

       # Determine how many children each job has
  parentAge = cTimes - bTimes
  numKids = rpois(n = length(parentAge), 
                  lambda = lambda*parentAge)
  
       # Determine the birth and completion times of the children
  mapply(function(n, min, max) {
            births = sort(runif(n, min, max))
            runtimes = rexp(n, rate = kappa)
            completes = rep(max, n) + runtimes
            data.frame(births, completes)
         },
         n = numKids , min = bTimes, max = cTimes,
         SIMPLIFY = FALSE)
}

# check the above work with a unit test 7.6
bTimes1 = 1:3
cTimes1 = c(3, 10, 15)

seed1 = 12062013
set.seed(seed1)

kids = genKidsV(bTimes1, cTimes1)
kids


kids2 = genKidsV(bTimes1, cTimes1)
sapply(kids2, nrow)

set.seed(seed1)
kids3 = genKidsV(bTimes = bTimes1, cTimes = cTimes1)
identical(kids, kids3)


# 7.7 Generate kids vectorized returning the ids of the kids and parents
genKidsV = function(bTimes, cTimes, parentID, lambda = 0.5, kappa = 0.3) {
       # Determine how many children each job has
  parentAge = cTimes - bTimes
  numKids = rpois(n = length(parentAge), lambda = lambda * parentAge)
  
  if (sum(numKids) == 0) return(NULL)
  
       # Determine the birth times of the children  
  kidStats = 
    mapply(function(n, min, max) {
      births = sort(runif(n, min, max))
      runtimes = rexp(n, rate = kappa)
      completes = rep(max, n) + runtimes
      data.frame(births, completes)
    },
           n = numKids , min = bTimes, max = cTimes, 
           SIMPLIFY = FALSE)
  
 
  return(data.frame(parentID = rep(parentID, numKids),
                    kidID = 1:sum(numKids), 
                    births = unlist(lapply(kidStats, "[[", "births")), 
                    completes = unlist(lapply(kidStats,"[[", "completes"))
                    ))
}

set.seed(seed1)
genKidsV(bTimes1, cTimes1, parentID = letters[1:3])
```

```{r include=FALSE, echo=FALSE}
familyTree = function(lambda = 0.5, kappa = 0.3, maxGen = 10) {
       # maxGen - maximum number of generations to observe
       # Return value - a list with 1 data frame per generation.
  allGens = vector(mode = "list", length = maxGen)
  
       # Generate the root of the tree
  allGens[[1]] = data.frame(parentID = NA, kidID = 1, births = 0,
                            completes = rexp(1, rate = kappa))
  
       # Generate future generations, one at a time.
  for (i in 2:maxGen) {
    nextGen = genKidsV(bTimes = allGens[[ (i - 1) ]]$births,
                       cTimes = allGens[[ (i - 1) ]]$completes,
                       parentID = allGens[[ (i - 1) ]]$kidID,
                       lambda = lambda, kappa = kappa)
    if (is.null(nextGen)) return(allGens[ 1:(i - 1) ])
    allGens[[ i ]] = nextGen
  }
  
  return(allGens)
}

set.seed(seed1)
tree = familyTree(lambda = 0.4, kappa = 1, maxGen = 10)

pdf("BA_FamilyTree.pdf", width = 8, height = 5)
oldPar = par(mar = c(4.1, 4.1, 0.5, 0.5))

set.seed(seed1)
g = familyTree(lambda = 0.4, kappa = 1, maxGen = 10)

maxLife = max(sapply(g, function(gen) max(gen$completes)))
numFamily = sum(sapply(g, nrow))

plot(0,0, ylim = c(0.5, numFamily + 0.5), xlim = c(0, maxLife),
     xlab = "Time", ylab = "", type ="n",
     axes = FALSE)

box()
axis(1)
numGen = length(g)
numKids = sapply(g, nrow)
treeN = g[ 2:(numGen + 1) ]
birthNum = c(0, cumsum(sapply(g, nrow))[ -length(g)])
axis(2, at = birthNum + 1, 
     labels = paste("Gen", 1:numGen), tick = FALSE, las = 1)

mapply(function(gen, nextGen, birthNum) {
  birthOrder = birthNum + (1:nrow(gen))
  segments(x0 = gen$births, x1 = gen$completes, y0 = birthOrder,
          lwd = 3, col = "grey")
  abline(h = 0.5 + max(birthOrder), lty = 2, col="black" )
  if (all(!is.na(nextGen$births)))
  points(x = nextGen$births, y = birthOrder[nextGen$parentID],
         pch = 4)
  },
  gen = g, nextGen = treeN, birthNum = birthNum )
par(oldPar)
dev.off()

seed2 = 12212013
set.seed(seed2)
tree = familyTree(lambda = 0.3, kappa = 0.5, maxGen = 10)

length(tree)

sapply(tree, nrow)

sum(sapply(tree, nrow))

set.seed(seed2)
tree = familyTree(lambda = 0.3, kappa = 0.5, maxGen = 15)
sapply(tree[ - (1:9) ], nrow)

set.seed(seed2)
tree = familyTree(lambda = 1, kappa = 0.5, maxGen = 10)

length(tree)

sapply(tree, nrow)

sum(sapply(tree, nrow))

tree = familyTree(lambda = 1, kappa = 0.5, maxGen = 10)
sum(sapply(tree, nrow))

sapply(tree, function(gen) range(gen$births))

sapply(tree, function(gen) range(gen$completes))

familyTreeT = function(lambda = 0.5, kappa = 0.3, maxTime = 8) {
       # maxTime - maximum length of time to observe the process
  allGens = list()
  
       # Generate the root of the 
  allGens[[1]] = data.frame(parentID = NA, kidID = 1, 
                            births = 0, 
                            completes = rexp(1, rate = kappa))
  
       # Generate the future generations, one at a time.
  numGens = 1
  while (TRUE) {
    nextGen = genKidsV(bTimes = allGens[[ numGens ]]$births,
                       cTimes = allGens[[ numGens ]]$completes,
                       parentID = allGens[[ numGens ]]$kidID,
                       lambda = lambda, kappa = kappa)
    
    if ( is.null(nextGen) | (min(nextGen$births) > maxTime) ) {
           # If complete after maxTime set complete to NA
      allGens = lapply(allGens, function(gen) {
        gen$completes[ gen$completes > maxTime ] = NA
        gen
      })
      return(allGens)
    }
    
    # Drop those born after maxTime 
    nextGen = nextGen[ nextGen$births <= maxTime , ]
    nextGen$kidID = 1:nrow(nextGen)
    numGens = numGens + 1
    allGens[[ numGens ]] = nextGen    
  }  
}	

familyTree = function(lambda = 0.5, kappa = 0.3, 
                      maxGen = 10, maxOffspring = 1000) {

       # Return value - a list with 1 data frame per generation.
  allGens = vector(mode = "list", length = maxGen)
  
       # Generate root of the tree
  allGens[[1]] = data.frame(parentID = NA, kidID = 1, 
                            births = 0, 
                            completes = rexp(1, rate = kappa))
  
  currentNumOffspring = 0

       # Generate future generations, one at a time.
  for (i in 2:maxGen) {
    nextGen = genKidsV(bTimes = allGens[[ (i - 1) ]]$births,
                       cTimes = allGens[[ (i - 1) ]]$completes,
                       parentID = allGens[[ (i - 1) ]]$kidID,
                       lambda = lambda, kappa = kappa)
    if (is.null(nextGen)) return(allGens[ 1:(i - 1) ]) 
    allGens[[ i ]] = nextGen
    currentNumOffspring = currentNumOffspring + nrow(nextGen)
    if (currentNumOffspring > maxOffspring) 
      return(allGens[1:i])
  }  
  allGens
}

set.seed(seed2)
tree = familyTree(lambda = 1, kappa = 0.5, 
                  maxGen = 100, maxOffspring = 1000)

length(tree)

sapply(tree, nrow)

sum(sapply(tree, nrow))

set.seed(seed2)
treeVT = familyTreeT(lambda = 1, kappa = 0.5)

numGen = 5
treeSub = treeVT[1:numGen]
numKids = sapply(treeSub, nrow)
numFamily = sum(numKids)

treeSub = lapply(treeSub, 
                 function(gen) {
                   gen$completes[ is.na(gen$completes) ] = 8
                   gen
                   })

treeSubN = treeVT[ 2:(numGen + 1) ]
treeSubN = lapply(treeSubN, 
                  function(gen) {
                    gen$completes[ is.na(gen$completes) ] = 8
                    gen
                 })
  
maxLife = max(sapply(treeSub, function(gen) max(gen$completes)))

birthNum = c(0, cumsum(numKids)[ -numGen ])

pdf("BA_FamilyTreeVT.pdf", width = 8, height = 15)
oldPar = par(mar = c(4.1, 4.1, 0.5, 0.5))

plot(0,0, ylim = c(0.5, numFamily), xlim = c(0, maxLife),
     xlab = "Time", ylab = "", type ="n",
     axes = FALSE)

box()
axis(1)
axis(2, at = birthNum + c(0.5, rep(1, (numGen-1))), 
     labels = paste("Gen", 1:numGen), tick = FALSE, las = 1)

mapply(function(gen, nextGen, birthNum) {
  birthOrder = birthNum + (1:nrow(gen))
  segments(x0 = gen$births, x1 = gen$completes, y0 = birthOrder,
           lwd = 3, col = "grey")
  abline(h = 0.5 + max(birthOrder), lty = 2, col="black" )
  points(x = nextGen$births, y = birthOrder[nextGen$parentID],
           pch = 4)
}, gen = treeSub, nextGen = treeSubN, birthNum = birthNum )

abline(v = maxLife)

par(oldPar)
dev.off()
```

```{r include=FALSE, echo=FALSE}
exptOne = function(l, k, mG, mO){
       # Helper function to call familyTree
       # Returns - summary statistics for analysis,
  
  aTree = familyTree(lambda = l, kappa = k, maxGen = mG,
                      maxOffspring = mO)
  numGen = length(aTree)
  numJobs = sum(sapply(aTree, nrow))
  c(numGen, numJobs)
}

set.seed(seed2)
exptOne(1, 0.5, 100, 1000)

familyTreeT = function(lambda = 0.5, kappa = 0.3, maxTime = 8) {
  # maxTime - maximum length of time to observe the family
  allGens = list()
  
  # Generate the root of the 
  allGens[[1]] = data.frame(parentID = NA, kidID = 1, 
                            births = 0, 
                            completes = rexp(1, rate = kappa))
  
  # Generate the future generations, one at a time.
  numGens = 1
  while (TRUE) {
    nextGen = genKidsV(bTimes = allGens[[ numGens ]]$births,
                       cTimes = allGens[[ numGens ]]$completes,
                       parentID = allGens[[ numGens ]]$kidID,
                       lambda = lambda, kappa = kappa)
    
    treeTerminated = is.null(nextGen)
    if ( !treeTerminated ) {
      birthsPastObsTime = min(nextGen$births) > maxTime
    }
    else birthsPastObsTime = TRUE
    
    if ( treeTerminated | birthsPastObsTime ) {
      # If complete after maxTime set complete to NA
      allGens = lapply(allGens, function(gen) {
        gen$completes[ gen$completes > maxTime ] = NA
        gen
      })
      return(allGens)
    }
    
    # Drop those born after maxTime 
    nextGen = nextGen[ nextGen$births <= maxTime , ]
    nextGen$kidID = 1:nrow(nextGen)
    numGens = numGens + 1
    allGens[[ numGens ]] = nextGen    
  }  
}

MCBA = function(params, repeats = 5, mG = 10, mO = 1000){
       # params: matrix columns of lambda and kappa values
       # For each lambda and kappa pair, run "repeats" times
  
  n = nrow(params)
  mcResults = vector("list", length = n)
  
  for (i in 1:n) {
    cat("param set is ", i, "\n")
    mcResults[[i]] = replicate(repeats,
                               exptOne(l = params[i, 1],
                                       k = params[i, 2],
                                       mG = mG, mO = mO))
  }
  mcResults
}

trialKappas = c(0.1, 10, 0.1, 10)
trialLambdas = c(0.1, 0.1, 10, 10)
trialParams = matrix(c(trialLambdas, trialKappas), ncol = 2)
mcTrialOutput = MCBA(params = trialParams, repeats = 40, 
                     mG = 200, mO = 100000)

save(mcTrialOutput, file = "mcTrialOutput.rda")

pdf("BA_ScatterPlotNumGenByNumKids.pdf", width = 10, height = 8)

oldPar = par(mfrow = c(2, 2), mar = c(3,3,1,1))

mapply(function(oneSet, lambda, kappa) {
  plot(x = oneSet[2,], y = jitter(oneSet[1, ], 1), log = "x",
       ylim = c(1,20), xlim = c(1, 10^7), pch = 19, cex = 0.6)
  text(x = 50, y = 15, bquote(paste(lambda == .(lambda))) )
  text(x = 300, y = 15, bquote(paste(kappa == .(kappa))) )
  }, 
  mcTrialOutput, lambda = trialLambdas, kappa = trialKappas)

par(oldPar)

dev.off()
```

```{r include=FALSE, echo=FALSE}
lambdas = c(seq(0.1, 0.6, by = 0.1), seq(0.8, 2, by = 0.2), 
            seq(2.25, 3, by = 0.25))
kappas = c(lambdas, 3.25, 3.50, 3.75, 4.00, 4.50, 5.00)

paramGrid = as.matrix(expand.grid(lambdas, kappas))

pdf("BA_Scatterplot3Dkids.pdf", width = 7, height = 6)

mcGrid = MCBA(params = paramGrid, repeats = 40, mG = 20,
                mO = 1000)

save(mcGrid, file = "mcGridOutput.rda")

logUQkids = sapply(mcGrid, function(x) 
               log(quantile(x[2, ], probs = 0.75), base = 10))

UQCut = cut(logUQkids, breaks = c(-0.1, 0.5, 2, max(logUQkids)) )
color3 = c("#b3cde3aa", "#8856a7aa", "#810f7caa")
colors = color3[UQCut]

library(scatterplot3d)
sdp = scatterplot3d(x = paramGrid[ , 1], y = paramGrid[ , 2], 
                    z = logUQkids, pch = 15, color = colors,
                    xlab = "Lambda", ylab = "Kappa",
                    zlab = "Upper Quartile Offspring",
                    angle = 120, type="h")

legend("left", inset = .08, bty = "n", cex = 0.8,
       legend = c("[0, 0.5)", "[0.5, 2)", "[2, 5)"), 
       fill = color3)

dev.off()
```

```{r include=FALSE, echo=FALSE}
pdf("BA_ImageMapAlive.pdf", width = 7, height = 7)
oldPar  = par(mar = c(4.1, 4.1, 0.5, 0.5))

mcGridAlive = sapply(mcGrid, function(oneParamSet) {
  sum((oneParamSet[1,] == 20) | (oneParamSet[2,] > 1000)) / 
    length(oneParamSet[2,]) })

filled.contour(lambdas, kappas, 
               matrix(mcGridAlive, nrow = length(lambdas), 
                      ncol = length(kappas)), 
               xlab = "Lambda", ylab = "Kappa", 
               xlim = c(0.1, 3), ylim = c(0.1, 3.1)) 

par(oldPar)
dev.off()
```

```{r include=FALSE, echo=FALSE}
pdf("BA_ImageMapAtleast20Kids.pdf", width = 7, height = 7)
oldPar  = par(mar = c(4.1, 4.1, 2, 1))

mcGridProp20kids = sapply(mcGrid, function(oneParamSet) {
  sum(oneParamSet[2,] > 19) / length(oneParamSet[2,]) })

mcGridProp20kidsMat = matrix(mcGridProp20kids, 
                             nrow = length(lambdas), 
                             ncol = length(kappas))

breaks = c(0, 0.10, 0.2, 0.3, 0.5, 0.7, 0.9, 1)
colors = rev(rainbow(10))[-(1:3)]

image(lambdas, kappas, mcGridProp20kidsMat, col = colors,
      breaks = breaks, xlab = "Lambda", ylab = "Kappa", 
      xlim = c(0.05, 3.05), ylim = c(0.05, 3.05))

midBreaks = (breaks[ -8 ] + breaks[ -1 ]) / 2
legend(x = 0.1, y = 3.25, legend = midBreaks, fill = colors, 
       bty = "n", ncol = 7,  xpd = TRUE)

par(oldPar)
dev.off()

```
Analysis for question 10
```{r  Plot 1, include=FALSE, echo=FALSE}
lambdas = c(seq(0.01, 0.2, by = 0.1), seq(0.4, 1.6, by = 0.2), 
            seq(1.25, 2, by = 0.25))
kappas = c(lambdas, 1.00, 1.00, 1.00, 1.00, 1.00, 1.00)

paramGrid = as.matrix(expand.grid(lambdas, kappas))

pdf("Plot 1.pdf", width = 7, height = 6)

mcGrid = MCBA(params = paramGrid, repeats = 40, mG = 20,
                mO = 1000)

save(mcGrid, file = "mcGridOutput.rda")

logUQkids = sapply(mcGrid, function(x) 
               log(quantile(x[2, ], probs = 0.75), base = 10))

UQCut = cut(logUQkids, breaks = c(-0.1, 0.5, 2, max(logUQkids)) )
color3 = c("#b3cde3aa", "#8856a7aa", "#810f7caa")
colors = color3[UQCut]

library(scatterplot3d)
sdp = scatterplot3d(x = paramGrid[ , 1], y = paramGrid[ , 2], 
                    z = logUQkids, pch = 15, color = colors,
                    xlab = "Lambda", ylab = "Kappa",
                    zlab = "Upper Quartile Offspring",
                    angle = 120, type="h")

legend("left", inset = .08, bty = "n", cex = 0.8,
       legend = c("[0, 0.5)", "[0.5, 2)", "[2, 5)"), 
       fill = color3)

dev.off()
```

```{r Plot 2, include = FALSE, echo = FALSE}
lambdas = c(seq(0.015, 0.4, by = 0.1), seq(0.6, 1.8, by = 0.2), 
            seq(1.75, 2.75, by = 0.25))
kappas = c(lambdas, 1.00, 1.00, 1.00, 1.00, 1.00, 1.00)

paramGrid = as.matrix(expand.grid(lambdas, kappas))

pdf("Plot 2.pdf", width = 7, height = 6)

mcGrid = MCBA(params = paramGrid, repeats = 40, mG = 20,
                mO = 1000)

save(mcGrid, file = "mcGridOutput.rda")

logUQkids = sapply(mcGrid, function(x) 
               log(quantile(x[2, ], probs = 0.75), base = 10))

UQCut = cut(logUQkids, breaks = c(-0.1, 0.5, 2, max(logUQkids)) )
color3 = c("#b3cde3aa", "#8856a7aa", "#810f7caa")
colors = color3[UQCut]

library(scatterplot3d)
sdp = scatterplot3d(x = paramGrid[ , 1], y = paramGrid[ , 2], 
                    z = logUQkids, pch = 15, color = colors,
                    xlab = "Lambda", ylab = "Kappa",
                    zlab = "Upper Quartile Offspring",
                    angle = 120, type="h")

legend("left", inset = .08, bty = "n", cex = 0.8,
       legend = c("[0, 0.5)", "[0.5, 2)", "[2, 5)"), 
       fill = color3)

dev.off()
```
```{r Plot 3, include = FALSE, echo = FALSE}
lambdas = c(seq(0.1, 0.6, by = 0.1), seq(0.8, 2.0, by = 0.2), 
            seq(2.25, 3.00, by = 0.25))
kappas = c(lambdas, 1.00, 1.00, 1.00, 1.00, 1.00, 1.00)

paramGrid = as.matrix(expand.grid(lambdas, kappas))

pdf("Plot 3.pdf", width = 7, height = 6)

mcGrid = MCBA(params = paramGrid, repeats = 40, mG = 20,
                mO = 1000)

save(mcGrid, file = "mcGridOutput.rda")

logUQkids = sapply(mcGrid, function(x) 
               log(quantile(x[2, ], probs = 0.75), base = 10))

UQCut = cut(logUQkids, breaks = c(-0.1, 0.5, 2, max(logUQkids)) )
color3 = c("#b3cde3aa", "#8856a7aa", "#810f7caa")
colors = color3[UQCut]

library(scatterplot3d)
sdp = scatterplot3d(x = paramGrid[ , 1], y = paramGrid[ , 2], 
                    z = logUQkids, pch = 15, color = colors,
                    xlab = "Lambda", ylab = "Kappa",
                    zlab = "Upper Quartile Offspring",
                    angle = 120, type="h")

legend("left", inset = .08, bty = "n", cex = 0.8,
       legend = c("[0, 0.5)", "[0.5, 2)", "[2, 5)"), 
       fill = color3)

dev.off()
```
```{r Plot 4, include= FALSE, echo= FALSE}
lambdas = c(seq(0.2, 0.8, by = 0.1), seq(1.0, 2.2, by = 0.2), 
            seq(2.50, 3.25, by = 0.25))
kappas = c(lambdas, 1.00, 1.00, 1.00, 1.00, 1.00, 1.00)

paramGrid = as.matrix(expand.grid(lambdas, kappas))

pdf("Plot 4.pdf", width = 7, height = 6)

mcGrid = MCBA(params = paramGrid, repeats = 40, mG = 20,
                mO = 1000)

save(mcGrid, file = "mcGridOutput.rda")

logUQkids = sapply(mcGrid, function(x) 
               log(quantile(x[2, ], probs = 0.75), base = 10))

UQCut = cut(logUQkids, breaks = c(-0.1, 0.5, 2, max(logUQkids)) )
color3 = c("#b3cde3aa", "#8856a7aa", "#810f7caa")
colors = color3[UQCut]

library(scatterplot3d)
sdp = scatterplot3d(x = paramGrid[ , 1], y = paramGrid[ , 2], 
                    z = logUQkids, pch = 15, color = colors,
                    xlab = "Lambda", ylab = "Kappa",
                    zlab = "Upper Quartile Offspring",
                    angle = 120, type="h")

legend("left", inset = .08, bty = "n", cex = 0.8,
       legend = c("[0, 0.5)", "[0.5, 2)", "[2, 5)"), 
       fill = color3)

dev.off()
```


```{r Plot 5, include= FALSE, echo=FALSE}
lambdas = c(seq(0.3, 0.9, by = 0.1), seq(1.6, 2.8, by = 0.2), 
            seq(3.00, 4.25, by = 0.25))
kappas = c(lambdas, 1.00, 1.00, 1.00, 1.00, 1.00, 1.00)

paramGrid = as.matrix(expand.grid(lambdas, kappas))

pdf("Plot 5.pdf", width = 7, height = 6)

mcGrid = MCBA(params = paramGrid, repeats = 40, mG = 20,
                mO = 1000)

save(mcGrid, file = "mcGridOutput.rda")

logUQkids = sapply(mcGrid, function(x) 
               log(quantile(x[2, ], probs = 0.75), base = 10))

UQCut = cut(logUQkids, breaks = c(-0.1, 0.5, 2, max(logUQkids)) )
color3 = c("#b3cde3aa", "#8856a7aa", "#810f7caa")
colors = color3[UQCut]

library(scatterplot3d)
sdp = scatterplot3d(x = paramGrid[ , 1], y = paramGrid[ , 2], 
                    z = logUQkids, pch = 15, color = colors,
                    xlab = "Lambda", ylab = "Kappa",
                    zlab = "Upper Quartile Offspring",
                    angle = 120, type="h")

legend("left", inset = .08, bty = "n", cex = 0.8,
       legend = c("[0, 0.5)", "[0.5, 2)", "[2, 5)"), 
       fill = color3)

dev.off()
```
#Analysis with Parameter Adjustment:

For case study 12 the project team decided to answer Question 10.  Parameters selected for the baseline study (plot 3) are from the example presented in the book used for the class.  The project team then decided to obtain sets of numbers for lambda that would be above and below the baseline study.  The selection, in theory, should reveal any possible trends in the data.  The diagram below is the parameter selection and the output.  

#Results:

For the case study, the randomly picked lambda values do reveal a possible trend in the data.  By keeping kappa, the lifetime value, constant and equal to one and increasing lambda numerically, the birth rates have a corresponding correlation with the lambda values. Starting with plot 1 (from above) the birth rate has the lowest density of the five trial runs.  The smallest overall values of lambda are in trial 1.  Revealed in the diagrams above as the lambda values change to higher values, there is a possible increase in birth rate seen in Plots 2 thru 5.  Plot 5 has highest overall lambda values and the highest density for birth rate in the study.   As Plot 1 is the most sparsely distributed output, one can see as you increase lambda there is an increase in density for the values in the Plots 1-5 for individual points in the diagram.    

#Conclusions:

After comparing the base study to the randomly selected lambda values for Question 10, it appears that there is a possible trend in the birth rates.  As lambda increases, the population density for birth rate has a corresponding result.  After researching the definition for lambda, one should possibly expect to see this change in birth rates.  By comparing the output from the Plots1 through Plot 5 above, one can see the possible trend in the increase of birth rates compared to the lambda values.  Because kappa, the lifetime of a job, is kept constant at one, the value changes in lambda truly depict or highlight birth rate changes.

